Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-11 inc/conclusions.tex
citations are now preceeded by non-breaking space (~)

In this thesis I have presented the extended Analytic Gaussian (AG) model, an effiecient numerical model based on the work of Michalik and Sipe \cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
For pulses containing $N$ carriers of charge $e$ and mass $m$, the model is simply these four equations,
\begin{subequations}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \label{eq:ag_model_gamma_conc} \,\text{.}
\end{gather}
\end{subequations}
where the subscript $\alpha$ denotes the directionality of the dynamics, either $\smallT$ for the transverse or $z$ for the longitudinal.
At any time $t$, the $\sigma_{\alpha}$ quantities characterize the spatial variance of the pulse and $\gamma_{\alpha}$ represent the momentum chirp across the pulse.
A third parameter, the local momentum variance $\eta_{\alpha}$, is also encoded in these equations.
By Liouville's theorem, which the AG model explicitly satisfies, we know that the product of the spatial variance $\sigma_{\alpha}$ and the local momentum variance $\eta_{\alpha}$ is constant throughout propagation (for perfect lensing and without clipping the puse), so we define the constant,
\begin{equation} \label{eq:define_big_gamma_conc}
  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} \,\text{.}
\end{equation}
By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc} we can get the extract these $\eta_{\alpha}$ parameters while only needing to compute the other four.
The parameter $\xi$ characterizes the ellipticity of the pulse,
\begin{equation}
  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}} \,\text{,}
\end{equation}
and the family of $L$ functions (defined in \ref{eq:ag_l_family}) are smooth and well behaved functions for all physically meaningful $\xi$.
Finally the $M_{\alpha}^{\lbrace f \rbrace}$ quantities define the $\alpha$ diretion influence of the $f^{\text{th}}$ external force.

@@

The Rose criterion \cite{rose_television_1948} states that in order to have proper gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera, the pulse must contain at least $10^8$ electrons.
We have proposed that the term ``ultrafast'' be used in the case where an electron pulse may be completely contained inside the acceleration region of the column.
The reason for this designation is that such pulses are subject to a different form of Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
\begin{equation}
  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d} \,\text{,}
\end{equation}
where $V_{DC}$ and $d$ are the accelerator voltage and length respectively.
In order to satisfy the both the Rose criterion and this ultrafast Child-Langmuir restriction, we expect that UEM instruments will need to employ a large-area emission source.
This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.} \cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
We have also designed and built a custom laser source capable of generating these electron pulses.

A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, which for the usual case of axially symmetric electron pulses has the form,
\begin{equation}
  \varepsilon_{\smallT} = \frac{1}{m c} \Delta x \Delta p_{\smallT} \,\text{,}
\end{equation}
where $\Delta x$ and $\Delta p_{\smallT}$ are the transverse spatial uncertainty and transverse momentum uncertainty, respectively.
This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones \cite{oshea_reversible_1998}.
The normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
Since we have already established that a large area emission source must be used, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.

@@

I have also presented our efforts to employ periodic nanostructured photocathode surfaces, designed to allow the incident laser to drive a surface plasmon, in hopes that a surface plasmon field might reduce the transverse emittance \cite{zawadzka_evanescent_2001,kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
Though we had success in driving the plasmon, we were unable to demonstrate reduction in $\Delta p_{\smallT}$.
We believe the problems were due to disturbances in the acceleration field near these nanostructures, though efforts to mitigate those disturbances has yet been unsuccessful.

>>>>

In this thesis I have presented the extended Analytic Gaussian (AG) model, an effiecient numerical model based on the work of Michalik and Sipe~\cite{michalik_analytic_2006}, but extended so as to be useful in modeling the dynamics of ultrafast electron pulses in microscope columns.
For pulses containing $N$ carriers of charge $e$ and mass $m$, the model is simply these four equations,
\begin{subequations}
\begin{gather}
  \frac{d\sigma_{\alpha}}{dt} = \frac{2\gamma_{\alpha}}{m} \\
  \frac{d\gamma_{\alpha}}{dt} = \frac{ \Gamma_{\alpha}^2 + \gamma_{\alpha}^2 }{\sigma_{\alpha} m}
    + \frac{N e^2}{4\pi\varepsilon_0} \frac{1}{6 \sqrt{\sigma_{\alpha}\pi}} L_{\alpha}(\xi) + \sigma_{\alpha} \sum\limits_{f} M_{\alpha}^{\lbrace f \rbrace} \label{eq:ag_model_gamma_conc} \,\text{.}
\end{gather}
\end{subequations}
where the subscript $\alpha$ denotes the directionality of the dynamics, either $\smallT$ for the transverse or $z$ for the longitudinal.
At any time $t$, the $\sigma_{\alpha}$ quantities characterize the spatial variance of the pulse and $\gamma_{\alpha}$ represent the momentum chirp across the pulse.
A third parameter, the local momentum variance $\eta_{\alpha}$, is also encoded in these equations.
By Liouville's theorem, which the AG model explicitly satisfies, we know that the product of the spatial variance $\sigma_{\alpha}$ and the local momentum variance $\eta_{\alpha}$ is constant throughout propagation (for perfect lensing and without clipping the puse), so we define the constant,
\begin{equation} \label{eq:define_big_gamma_conc}
  \Gamma_{\alpha}^2 \equiv \sigma_{\alpha} \eta_{\alpha} \,\text{.}
\end{equation}
By evaluating this constant once, presumably from initial conditions, and later applying \ref{eq:define_big_gamma_conc} we can get the extract these $\eta_{\alpha}$ parameters while only needing to compute the other four.
The parameter $\xi$ characterizes the ellipticity of the pulse,
\begin{equation}
  \xi \equiv \sqrt{\dfrac{\sigma_{z}}{\sigma_{\smallT}}} \,\text{,}
\end{equation}
and the family of $L$ functions (defined in \ref{eq:ag_l_family}) are smooth and well behaved functions for all physically meaningful $\xi$.
Finally the $M_{\alpha}^{\lbrace f \rbrace}$ quantities define the $\alpha$ diretion influence of the $f^{\text{th}}$ external force.

@@

The Rose criterion~\cite{rose_television_1948} states that in order to have proper gray-scaling in a single-shot DTEM employing a 1k X 1k CCD camera, the pulse must contain at least $10^8$ electrons.
We have proposed that the term ``ultrafast'' be used in the case where an electron pulse may be completely contained inside the acceleration region of the column.
The reason for this designation is that such pulses are subject to a different form of Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923,valfells_effects_2002}, which limits its the number of electrons $N_{crit}$ that may be generated by a laser of HW1/eM width $w$ before the pulse is subject to deleterious distortions,
\begin{equation}
  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d} \,\text{,}
\end{equation}
where $V_{DC}$ and $d$ are the accelerator voltage and length respectively.
In order to satisfy the both the Rose criterion and this ultrafast Child-Langmuir restriction, we expect that UEM instruments will need to employ a large-area emission source.
This has necessitated building a custom acceleration region based on the design of Togawa \textit{et al.}~\cite{togawa_ceb6_2007} which boasts a moderately uniform acceleration field near the axis of the accelerator, as well as other large-aperture electron-optical elements for our prototype UEM column.
We have also designed and built a custom laser source capable of generating these electron pulses.

A common figure of merit for electron beams is its normalized transverse emittance~\cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, which for the usual case of axially symmetric electron pulses has the form,
\begin{equation}
  \varepsilon_{\smallT} = \frac{1}{m c} \Delta x \Delta p_{\smallT} \,\text{,}
\end{equation}
where $\Delta x$ and $\Delta p_{\smallT}$ are the transverse spatial uncertainty and transverse momentum uncertainty, respectively.
This quantity is conserved under propagation through free space and perfect electron-optical, though it may increase for imperfect ones~\cite{oshea_reversible_1998}.
The normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
Since we have already established that a large area emission source must be used, the only avenue to reduce $\varepsilon_{\smallT}$ is to reduce the transverse momentum uncertainty, which is dependent on the photocathode and photoemission process employed.

@@

I have also presented our efforts to employ periodic nanostructured photocathode surfaces, designed to allow the incident laser to drive a surface plasmon, in hopes that a surface plasmon field might reduce the transverse emittance~\cite{zawadzka_evanescent_2001,kupersztych_ponderomotive_2001,kupersztych_anomalous_2005,li_surface_2013}.
Though we had success in driving the plasmon, we were unable to demonstrate reduction in $\Delta p_{\smallT}$.
We believe the problems were due to disturbances in the acceleration field near these nanostructures, though efforts to mitigate those disturbances has yet been unsuccessful.

### 2013-04-11 inc/design_considerations/considerations.tex
citations are now preceeded by non-breaking space (~)

It is straightforward to show that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
The Rose criterion \cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
For a 1k $\times$ 1k CCD camera, this implies an incident electron pulse with $N=10^8$ \cite{armstrong_prospects_2007}.

@@

It should be noted that for diffraction studies performed in a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse \cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.

@@

Since we are now aware that there is a large minimum number of electrons needed per pulse, it is only natural to ask if there is the practical limit on how much charge (i.e., the number of electrons/pulse, $N$) can be extracted from a pulsed laser-driven DC photoelectron gun.
When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}} \approx 0.1-1$ns, typically, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law \cite{child_discharge_1911,langmuir_effect_1923}:
\begin{equation}
  J_{CL} = \frac{4 \varepsilon_0}{9} \sqrt{\frac{2e}{m}} \frac{V_{DC}^{3/2}}{d^2} \,\text{,}
\end{equation}
where $e$ is the electron charge and $\varepsilon_0$ is the vacuum permittivity.

@@

In this short pulse case, an equivalent diode approximation \cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
\begin{equation}
  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL} \,\text{,}
\end{equation}
where $\chi = \tau_e / t_{gun}$.
When $\chi \ll 1$, the usual case in UEM, the critical current density is simply given by
\begin{equation}
  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e} \,\text{,}
\end{equation}
where the DC gun field $E_{DC} = V_{DC}/d$.
This relation may be rewritten to yield a critical number of photoemitted electrons/pulse before virtual cathode formation:
\begin{equation} \label{eq:n_crit}
  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d} \,\text{.}
\end{equation}
In obtaining the above equation, we have used an irradiated photocathode area of $\pi w^2$ associated with an incident Gaussian laser pulse of the form of \ref{eq:laser_form}.
The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected, because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.

For a typical DC gun with $E_{DC} \approx $ 30 kV/cm (e.g., $V_{DC} = $ 100 kV and $d \approx $ 3 cm), \ref{eq:n_crit} reveals that the generation of $10^8$ electrons/pulse needed for single-shot imaging applications in a DTEM with a 1k $\times$ 1k CCD detector requires an incident laser beam area $\pi w^2$ of about 1 mm$^2$ --- the virtual cathode being formed for the extraction of more than $1.7 \times 10^8$ electrons/mm$^2$ in this short pulse regime.
We note that \ref{eq:n_crit} does not express a limit on the number of electrons/pulse $N$ that can be emitted from a pulsed laser-driven DC photoelectron gun, it merely indicates a critical number before virtual cathode formation occurs.
More electrons can be produced, at the expense of a severely distorted temporal electron pulse profile \cite{valfells_effects_2002}.

@@

A common figure of merit for electron beams is its normalized transverse emittance \cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
\begin{equation}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > - \left < x p_{\smallT} \right >^2 } \,\text{,}
\end{equation}
where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_{\smallT}$ are the transverse position and momentum respectively.
For the usual case of axially symmetric electron pulses, we may use the simplified form
\begin{equation} \label{eq:emittance_definition}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > } \,\text{.}
\end{equation}
By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throughout pulse propagation (for free-space and perfect electron-optical elements), and therefore is related directly to the initial photoemission properties of the system.
For a laser pulse of the form of \ref{eq:laser_form} we have simply
\begin{equation} \label{eq:emittance_definition_w}
  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_{\smallT}^2 \right >}{2} } \,\text{.}
\end{equation}
Note that imperfect electron-optical elements may cause ``emittance growth'' \cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.

This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred \cite{berger_dc_2009}.
Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
Though deriving an absolute limit on the allowable transverse emittance for imaging in a UEM would depend on many column-specific factors, since diffraction is expected to impose less stringent requirements on the beam than imaging, we may consider it when searching for an upper bound on transverse emittance.

In Ref. \cite{berger_dc_2009}, we explore the impact of these limitations for an ultrafast electron diffraction (UED) experiment.
We find that for a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, one may only resolve a fractional change $\Delta a / a$ in a crystalline material's Bragg plane spacing $a$ in the $M^{th}$ diffraction order if
\begin{equation}
  \varepsilon_{\smallT} \le \frac{5 \hbar}{m c} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
\end{equation}
However, because of lower limit that exists on $w$ from \ref{eq:minimum_w}, we can say that these requirements impose an upper limit on the initial transverse momentum spread
\begin{equation}
  \sqrt{ \left < p_{\smallT}^2 \right > } = \sqrt{ \eta_{\smallT} (0) } \le \frac{5\sqrt{2} \hbar}{w} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
\end{equation}

@@

One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, thereby reducing the excess photoemission energy $\Delta E$.
Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.
In Chapter \ref{chap:photocathode}, I will explore several more exotic photoemission processes in an attempt to reduce $\varepsilon_{\smallT}$ without this penalty.

>>>>

It is straightforward to show that in order to make a useful image in an electron microcope, the number of electrons collected must be above some minimum number.
The Rose criterion~\cite{rose_television_1948} indicates that an average detector signal strength of $\sim$100 electrons/pixel is required for adequate gray scaling of the image.
For a 1k $\times$ 1k CCD camera, this implies an incident electron pulse with $N=10^8$~\cite{armstrong_prospects_2007}.

@@

It should be noted that for diffraction studies performed in a DTEM, many fewer electrons per pulse are required for single-shot experiments; typically, about $10^6$ electrons/pulse~\cite{armstrong_practical_2007} as significantly less signal is required to determine the position of discrete single-crystal Bragg diffraction spots or diffraction rings for polycrystalline specimens.

@@

Since we are now aware that there is a large minimum number of electrons needed per pulse, it is only natural to ask if there is the practical limit on how much charge (i.e., the number of electrons/pulse, $N$) can be extracted from a pulsed laser-driven DC photoelectron gun.
When the incident laser pulse duration $\tau$ is longer than the time-of-flight (TOF) $t_{gun}$ of a single electron across the acceleration gap $d$ of the DC gun with a potential $V_{DC}$, $\tau > t_{gun} = d \sqrt{2m/eV_{DC}} \approx 0.1-1$ns, typically, then the maximum photocurrent density $J_{CL}$ is given by the well-known Child-Langmuir law~\cite{child_discharge_1911,langmuir_effect_1923}:
\begin{equation}
  J_{CL} = \frac{4 \varepsilon_0}{9} \sqrt{\frac{2e}{m}} \frac{V_{DC}^{3/2}}{d^2} \,\text{,}
\end{equation}
where $e$ is the electron charge and $\varepsilon_0$ is the vacuum permittivity.

@@

In this short pulse case, an equivalent diode approximation~\cite{valfells_effects_2002} allows for the definition of a critical current density associated with virtual cathode formation in terms of the steady-state Child-Langmuir limit:
\begin{equation}
  J_{crit} = \frac{27}{4} \left ( \frac{ 1 - \sqrt{1-\tfrac{1}{3} \chi^2} }{ \chi^3 } \right ) J_{CL} \,\text{,}
\end{equation}
where $\chi = \tau_e / t_{gun}$.
When $\chi \ll 1$, the usual case in UEM, the critical current density is simply given by
\begin{equation}
  J_{crit} = \frac{\varepsilon_0 E_{DC}}{\tau_e} \,\text{,}
\end{equation}
where the DC gun field $E_{DC} = V_{DC}/d$.
This relation may be rewritten to yield a critical number of photoemitted electrons/pulse before virtual cathode formation:
\begin{equation} \label{eq:n_crit}
  N_{crit} = \frac{\pi \varepsilon_0 w^2 V_{DC}}{e d} \,\text{.}
\end{equation}
In obtaining the above equation, we have used an irradiated photocathode area of $\pi w^2$ associated with an incident Gaussian laser pulse of the form of \ref{eq:laser_form}.
The relation clearly indicates that $N_{crit}$ is only dependent upon the initial DC surface field (i.e., $E_{DC}$) at the photocathode and the illuminated area --- as expected, because a virtual cathode forms when the initial photocathode surface charge density associated with $E_{DC}$ is photoemitted, effectively screening the cathode from the acceleration field.

For a typical DC gun with $E_{DC} \approx $ 30 kV/cm (e.g., $V_{DC} = $ 100 kV and $d \approx $ 3 cm), \ref{eq:n_crit} reveals that the generation of $10^8$ electrons/pulse needed for single-shot imaging applications in a DTEM with a 1k $\times$ 1k CCD detector requires an incident laser beam area $\pi w^2$ of about 1 mm$^2$ --- the virtual cathode being formed for the extraction of more than $1.7 \times 10^8$ electrons/mm$^2$ in this short pulse regime.
We note that \ref{eq:n_crit} does not express a limit on the number of electrons/pulse $N$ that can be emitted from a pulsed laser-driven DC photoelectron gun, it merely indicates a critical number before virtual cathode formation occurs.
More electrons can be produced, at the expense of a severely distorted temporal electron pulse profile~\cite{valfells_effects_2002}.

@@

A common figure of merit for electron beams is its normalized transverse emittance~\cite{jensen_emittance_2010}, $\varepsilon_{\smallT}$, defined as 
\begin{equation}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > - \left < x p_{\smallT} \right >^2 } \,\text{,}
\end{equation}
where $m$ is the electron mass, $c$ is the speed of light in vacuum and $x$ and $p_{\smallT}$ are the transverse position and momentum respectively.
For the usual case of axially symmetric electron pulses, we may use the simplified form
\begin{equation} \label{eq:emittance_definition}
  \varepsilon_{\smallT} = \frac{1}{m c} \sqrt{ \left < x^2 \right > \left < p_{\smallT}^2 \right > } \,\text{.}
\end{equation}
By Liouville's Theorem (Section \ref{sec:liouville}), we know that this quantity is conserved throughout pulse propagation (for free-space and perfect electron-optical elements), and therefore is related directly to the initial photoemission properties of the system.
For a laser pulse of the form of \ref{eq:laser_form} we have simply
\begin{equation} \label{eq:emittance_definition_w}
  \varepsilon_{\smallT} = \frac{w}{m c} \sqrt{ \frac{\left < p_{\smallT}^2 \right >}{2} } \,\text{.}
\end{equation}
Note that imperfect electron-optical elements may cause ``emittance growth''~\cite{oshea_reversible_1998}, increasing the pulse's emittance after these elements.

This normalized transverse emittance limits the resolving capablility of the beam, so this quantity can be thought of as a measure of the beam quality, where a lower number is preferred~\cite{berger_dc_2009}.
Since \ref{eq:minimum_w} specifies a lower bound on the emission spot size $w$, knowing a UEM will exhibit higher resolution by employing an electron source with the minimum possible transverse emittance directly implies that every effort must be made to reduce the initial transverse momentum spread.
Though deriving an absolute limit on the allowable transverse emittance for imaging in a UEM would depend on many column-specific factors, since diffraction is expected to impose less stringent requirements on the beam than imaging, we may consider it when searching for an upper bound on transverse emittance.

In Ref.~\cite{berger_dc_2009}, we explore the impact of these limitations for an ultrafast electron diffraction (UED) experiment.
We find that for a desired spatial resolution per pixel $\Delta X$ (related to the magnification) using a $N_p$ x $N_p$ detector, one may only resolve a fractional change $\Delta a / a$ in a crystalline material's Bragg plane spacing $a$ in the $M^{th}$ diffraction order if
\begin{equation}
  \varepsilon_{\smallT} \le \frac{5 \hbar}{m c} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
\end{equation}
However, because of lower limit that exists on $w$ from \ref{eq:minimum_w}, we can say that these requirements impose an upper limit on the initial transverse momentum spread
\begin{equation}
  \sqrt{ \left < p_{\smallT}^2 \right > } = \sqrt{ \eta_{\smallT} (0) } \le \frac{5\sqrt{2} \hbar}{w} N_p \, \Delta X \frac{M}{a} \left ( \frac{\Delta a}{a} \right ) \,\text{.}
\end{equation}

@@

One might think to narrow this discrepancy by simply better matching the photocathode work function and the laser photon energy, thereby reducing the excess photoemission energy $\Delta E$.
Unfortunately, since the photoemission efficiency is proportional to $\Delta E^2$~\cite{shalaev_electron_1994}, though this would indeed reduce the transverse emittance, it would do so at the cost of a much reduced pulse charge.
In Chapter \ref{chap:photocathode}, I will explore several more exotic photoemission processes in an attempt to reduce $\varepsilon_{\smallT}$ without this penalty.

### 2013-04-11 inc/extension/external_forces/formal.tex
citations are now preceeded by non-breaking space (~)

Simulation of electron pulse propagation in UEM and modern UED apparatus \cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen. 
The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities \cite{oudheusden_electron_2007,veisz_hybrid_2007}.

@@

While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics \cite{michalik_analytic_2006}.

@@

The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations \cite{michalik_analytic_2006}.
As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled \cite{oshea_reversible_1998}.
The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.

@@

Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes \cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid \cite{luiten_how_2004}.
This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
Consequently, the AG approach is applicable to both the single electron per pulse limit \cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate \cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.

>>>>

Simulation of electron pulse propagation in UEM and modern UED apparatus~\cite{siwick_ultrafast_2002,oudheusden_electron_2007,reed_femtosecond_2006} requires the ability to model electron-optical elements designed to control the three-dimensional space-charge and dispersion induced broadening of the pulse, and hence deliver electron pulses of the desired quality to the specimen. 
The Analytic Gaussian model as presented by Michalik and Sipe (see Section \ref{sec:ms_model}) is only useful for modeling free-space propagation.
This is because there is no allowance for external forces, such as those imparted by the ``optical'' elements of an electron microscope column.
In this section, I show that the AG model can be extended to include external forces on the pulse, such as the influence of magnetic electron lenses and RF pulse compression cavities~\cite{oudheusden_electron_2007,veisz_hybrid_2007}.

@@

While the free-space AG model does not contain any external forces, it does include an internal force: the Coulomb force is included to model the repulsion of the pulse's electrons from each other.
As this force is self-influencing, it is the major souce of complexity in the model; indeed without this term, a complete analytic solutions is possible, which is identical to ``Gaussian'' optics~\cite{michalik_analytic_2006}.

@@

The presented extension to the AG model to include external forces acting on the electron pulse is valid only within the limits of the analytical method itself; in particular, its mean internal space-charge field and self-similar Gaussian approximations~\cite{michalik_analytic_2006}.
As a result, we emphasize that the presented AG model extension is valid only for the action of linear forces (i.e., forces that have a linear dependence on distance in the pulse frame) on the electron pulse due to the model's self-similar Gaussian approximation.
Since nonlinear forces are not considered, ``emittance growth'' effects arising from imperfect lenses cannot be modeled~\cite{oshea_reversible_1998}.
The inclusion of the action of nonlinear forces (e.g., due to higher-order space-charge effects, spherical magnetic lens aberrations, etc.) will require the use of more sophisticated electron pulse propagation models such as full particle tracking simulations.

@@

Nonetheless, for free-space propagation, the AG model of charge bunch dynamics has already been shown to be very consistent with full Monte Carlo (i.e., particle tracking) simulations for a wide variety of electron pulse shapes~\cite{michalik_analytic_2006,michalik_evolution_2009}, including the uniform ellipsoid~\cite{luiten_how_2004}.
This successful benchmarking is due primarily to the versatility of the AG model which results from its use of transverse and longitudinal pulse position and momentum variances.
Consequently, the AG approach is applicable to both the single electron per pulse limit~\cite{lobastov_four-dimensional_2005}, where momentum variances determine the pulse evolution and the model is exact (obeying Gaussian optics), and the high charge density limit in which space-charge effects dominate~\cite{luiten_how_2004,siwick_ultrafast_2002,cao_femtosecond_2003}.
It is this versatility combined with its computational efficiency that makes the presented extended AG model particularly suitable for rapid initial assessments of pulsed electron microscope column designs and electron pulse delivery systems in UED experiments.
Verification of the validity (and determination of the limits) of the extended AG model will, of course, require future comparison with both experiment and more complete simulations of electron pulse propagation dynamics (e.g., full particle tracking models) that include nonlinear forces, for both the intra-pulse space-charge field and the description of aberrations in electron optics.
